---
title: "ST558 Final Project"
author: "Yinzhou Zhu"
output: 
  github_document:
    toc: true
---
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Purpose of this project

# Data used in this project
Data is obtained by registering at The Surveillance, Epidemiology, and End Results ([NIH SEER](https://seer.cancer.gov/data/access.html)) and requesting access through SEER*Stat software. Tables are queried through the software and imported to R using `SEER2R` package.  

## SEER 18 data set
Data set chosen for this project is the [**SEER 18**](https://seer.cancer.gov/registries/terms.html), which contains data for cases from year 2000 through 2017. Although this set does not include data from earlier times such as from year 1975 or 1992 onward like the **SEER 9/13** data set, improvement in modern diagnosis, treatments and post-treatment care could impact the trend of survival. Thus, the **SEER 18** data set can be more appropriate for traditional ML predictions.  

## Data subsetting
I have removed all rows with "Unknown" in `Grade` variable, since cell type is an important aspect (regardless of whether the models eventually use this variable or not).  
In addition, only the most recent cases are used (2000-2010) due to the size of the database being really large. Doing this would also circumvent the issue where case is too recent and survival month data is not complete (cases where person is highly likely to succumb to cancer be currently still alive).  
User will be able to determine the ration between training and testing data.  

## Variables used
The Following variables are used from the data set:  
```{r variable table, echo=FALSE, results='asis'}
manual_table <- "
| Variables                        | Descriptions                                                                     |
|----------------------------------|:--------------------------------------------------------------------------------:|
| Sex                              | Includes 1= Male and 2=Female                                                    |
| Year of diagnosis                | Values are 2000-2019                                                             |
| Race Recode (W, B, AI, API)      | White,Black,American Indian/Alaska Native,Asian or Pacific Islander,Unknown      |
| Age Recode with <1 year olds     | Age of diagnosis:< 1 year, 1-4 years,..., 85+ years                              |
| Primary Site                     | Cecum, Lips etc. but with a code                                                 |
| Histological Type                | Microscopic composition of cells and/or tissue for primary tumor (code)          |
| Grade                            | Cell type identified                                                             |
| Laterality                       | The side of body/a paired organ on which the reportable tumor originated         |
| RX SUMM SURG PRIM SITE (1998+)   | Surgery of Primary Site                                                          |
| Survival months                  | Number of months survived                                                        |
"
cat(manual_table) 
```
